Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-09-02 paper/paper.tex
Advanced box moving skills
More writing

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs \citep{CP2016MCSPaper}. This is
largely because strong inference, based upon the degrees of vertices
\citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted. In
this work we discuss an intermediate problem, where we must map all but $k$ vertices of the pattern
graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then weakened
forms of degree and path based filterings are still effective in pruning the search space and
providing additional constraints respectively. We then show that combining these techniques does
lead to a practical algorithm which can scale to work with the families of graphs commonly used to
benchmark subgraph isomorphism algorithms: depending upon the benchmark family, in a substantial
portion of ; in many more cases, we can at least obtain an upper bound.

@@

For subgraph isomorphism, we can also do domain filtering at the top of search. We discuss this in
??. Additional constraints, discuss this in ??.

@@

\section{Domain Filtering using Degrees}

@@

\begin{proposition}
    \label{prop:deg}
    Let $p$ be a vertex in $P$ and $t$ a vertex in $T$. For both non-induced and induced
    $k$-less-subgraph isomorphisms, if $p \lessmap{k} t$ then
    $\deg(p) - k \le \deg(t)$.
\end{proposition}
\begin{proof}
Let $p$ be a vertex in $P$, and $t$ a vertex in $T$, with $p\mapsto t$. Then by the definition of subgraph isomorphisms, $\deg(p) \le \deg(t)$. Let $P'$ be $P$ less $k$ vertices and $p$ a vertex in $P'$. Then
\[
\deg_{P}(p)-k \le \deg_{P'}(p) \le \deg_{P}(p) \le \deg(t).
\]
\end{proof}

@@

\section{Filtering During Search Using Paths}

@@

If we remove up to $k$ vertices from the neighbourhood of $p$, it will impact the intersection of the neighbourhoods of $p$ and $q$, and by Proposition~\ref{prop:deg}, as $p\lessmap{k}t$
\[
\left| \N(p)\cap \N(q)\right| - k \leq deg(p) - k \leq deg(t).
\]
As $\left|\N(p)\cap \N(q)\right|\leq deg(p) \leq \left|\N(t)\cap \N(u)\right|\leq deg(t)$, we have
\[
\left|\N(p)\cap \N(q)\right| - k \leq deg(p) - k\leq \left|\N(t)\cap \N(u)\right|\leq deg(t)
\]
which is
\[
\paths(p, q, 2) - k \leq \paths(t, u, 2).
\]
\end{proof}

>>>>

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. This is
largely because strong inference, based upon the degrees of vertices
\citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted. In
this work we discuss an intermediate problem, where we must map all but $k$ vertices of the pattern
graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then weakened
forms of degree and path based filterings are still effective in pruning the search space and
providing additional constraints respectively. We then show that combining these techniques does
lead to a practical algorithm which can scale to work with the families of graphs commonly used to
benchmark subgraph isomorphism algorithms: depending upon the benchmark family, in a substantial
portion of ; in many more cases, we can at least obtain an upper bound.

@@

For subgraph isomorphism, we can also do domain filtering at the top of search. We discuss this in
\cref{section:degreefiltering}. Additional constraints, discuss this in \cref{section:pathfiltering}.

@@

We perform our experiments on systems with dual Intel Xeon CPU E5-2640 processors with 64GBytes of
RAM, running Ubuntu 14.04. Our algorithm was implemented in C++\footnote{URL removed for anonymous
review}, and source code was compiled using
GCC 5.3.0. We used a timeout of 1,000 seconds for each instance.

For datasets, we use the same 5,725 instances used in a recent work on portfolios of algorithms for
subgraph isomorphism \cite{thelionpaper}. This collection includes randomly generated scale-free
graphs \cite{constraints10}, an assortment of real-world graphs of varying sizes \cite{LV02},
randomly generated graph pairs (using bounded degree, regular mesh, and uniform models; all are
satisfiable), segmented images~\cite{pr15,cviu11}, meshes from modelling 3D objects \cite{cviu11},
and graphs close to the satisfiable-unsatisfiable phase transition
\cite{DBLP:conf/ijcai/McCreeshPT16}; all these instances are available in a simple text
format\footnote{\texttt{http://liris.cnrs.fr/csolnon/SIP.html}}.

?? Note about the size of these instances, and how they're much larger than MCS instances.

\section{Domain Filtering using Degrees}\label{section:degreefiltering}

@@

\begin{proposition}
    \label{prop:deg}
    Let $p$ be a vertex in $P$ and $t$ a vertex in $T$. For both non-induced and induced
    $k$-less-subgraph isomorphisms, if $p \lessmap{k} t$ then
    $\deg(p) - k \le \deg(t)$.
\end{proposition}
\begin{proof}
Let $p$ be a vertex in $P$, and $t$ a vertex in $T$, with $p\mapsto t$. Then by the definition of subgraph isomorphisms, $\deg(p) \le \deg(t)$. Let $P'$ be $P$ less $k$ vertices and $p$ a vertex in $P'$. Then
\[
\deg_{P}(p)-k \le \deg_{P'}(p) \le \deg_{P}(p) \le \deg(t). \qedhere
\]
\end{proof}

@@

\section{Filtering During Search Using Paths}\label{section:pathfiltering}

@@

If we remove up to $k$ vertices from the neighbourhood of $p$, it will impact the intersection of the neighbourhoods of $p$ and $q$, and by Proposition~\ref{prop:deg}, as $p\lessmap{k}t$
\[
\left| \N(p)\cap \N(q)\right| - k \leq deg(p) - k \leq deg(t).
\]
As $\left|\N(p)\cap \N(q)\right|\leq deg(p) \leq \left|\N(t)\cap \N(u)\right|\leq deg(t)$, we have
\[
\left|\N(p)\cap \N(q)\right| - k \leq deg(p) - k\leq \left|\N(t)\cap \N(u)\right|\leq deg(t)
\]
which is
\[
\paths(p, q, 2) - k \leq \paths(t, u, 2). \qedhere
\]
\end{proof}
